Ecosystem carbon balance in the Hawaiian Islands under different scenarios of future climate and land use change

### 2021-01-21 Manuscript/HI_Model_ms.Rmd
updated manuscript files

## Introduction
Terrestrial ecosystems are a major sink for atmospheric carbon dioxide (CO~2~), removing ~30% of human emissions on an annual basis and reducing the rate of increase in atmospheric CO~2~ [@keenan_terrestrial_2018; @friedlingstein_global_2019]. There is increasing recognition among policymakers that natural and angricultural ecosystems can contribute to climate mitigation, which has given rise to the popularity of “natural carbon solutions” [@cameron_ecosystem_2017]. Defined as conservation and land management efforts aimed at enhancing ecosystem carbon storage [@griscom_natural_2017], natural climate solutions are appealing because they are seen as cost-effective and readily available [@averchenkova_climate_2017; @cameron_ecosystem_2017; @fargione_natural_2018]. However, effective implementation is complicated by the uncertainty surrounding the future direction and magnitude of the land carbon sink, especially at the regional scale. Despite this uncertainty, evidence indicates that both interannual and long-term variability in carbon uptake by land ecosystems is driven primarily by fluctuations in climate, land use, and land cover change [@ahlstrom_dominant_2015; @prestele_current_2017; @friedlingstein_global_2019]. Spatially explicit future projections of ecosystem carbon balance incorporating the interactive effects of land use and climate could therefore provide a reference point to evaluate the effectiveness of land-based mitigation. Although a complex challenge, the growing number of sub-national jurisdictions that plan to incorporate land-based mitigation strategies into their emissions reduction efforts would benefit from understanding how future climate-biosphere feedbacks will affect ecosystem carbon balance in their respective regions [@sleeter_effects_2019].

The State of Hawaiʻi exemplifies the challenges of projecting future interactive effects of climate and land use change on ecosystem carbon balance. The main Hawaiian Islands are a complex mosaic of natural and human-dominated landscapes overlain by steep climate gradients across relatively short distances. Mean annual temperature ranges from ~4-24° C [@giambelluca_evapotranspiration_2014] and mean annual rainfall ranges from ~180-9500 mm [@giambelluca_online_2013], with minimal seasonal variation in climate. Temperatures have risen rapidly in the Hawaiian Islands since the mid 1970's [@giambelluca_secular_2008] and a long-term drying trend has persisted since the early 1920s [@frazier_spatial_2017], resulting in reduced forest biomass and productivity [@barbosa_effects_2017]. These same drying and warming trends have increased the frequency and intensity of wildland fire [@trauernicht_contemporary_2015; @trauernicht_vegetationrainfall_2019] with predictable negative effects on ecosystem carbon balance [@selmants_baseline_2017]. Ecosystem carbon stocks across the main Hawaiian Islands have also been strongly influenced by the legacy of past land-use change [@osher_effect_2003; @asner_high-resolution_2011]. Thousands of hectares of land were deforested beginning in the late 19th century to clear land for sugar plantations and cattle pasture [@cuddihy_alteration_1990]. Beginning in the latter half of the 20th century, this agricultural land has been steadily converted to urban areas, commercial forestry plantations, or simply abandoned and colonized by non-native grass species [@suryanata_diversified_2009; @perroy_evolving_2016].

Hawaiʻi was the first U.S. state to enact legislation supporting commitments and goals of the Paris climate agreement. Specifically, Hawaiʻi passed Act

>>>>

## Introduction
Terrestrial ecosystems are a major sink for atmospheric carbon dioxide (CO~2~), removing ~30% of human emissions on an annual basis and reducing the rate of increase in atmospheric CO~2~ [@keenan_terrestrial_2018; @friedlingstein_global_2019]. There is increasing recognition among policymakers that natural and angricultural ecosystems can contribute to climate mitigation, which has given rise to the popularity of “natural carbon solutions” [@cameron_ecosystem_2017]. Defined as conservation and land management efforts aimed at enhancing ecosystem carbon storage [@griscom_natural_2017], natural climate solutions are appealing because they are seen as cost-effective and readily available [@averchenkova_climate_2017; @cameron_ecosystem_2017; @fargione_natural_2018]. However, effective implementation is complicated by the uncertainty surrounding the future direction and magnitude of the land carbon sink, especially at the regional scale. Despite this uncertainty, evidence indicates that both interannual and long-term variability in carbon uptake by land ecosystems is driven primarily by fluctuations in climate, land use, and land cover change [@ahlstrom_dominant_2015; @prestele_current_2017; @friedlingstein_global_2019]. Incorporating the interactive effects of land use and climate into spatially explicit future projections of ecosystem carbon balance could therefore provide a reference point to evaluate the effectiveness of land-based mitigation. Although a complex challenge, the growing number of sub-national jurisdictions that plan to incorporate land-based mitigation strategies into their emissions reduction efforts would benefit from understanding how future climate-biosphere feedbacks will affect ecosystem carbon balance in their respective regions [@sleeter_effects_2019].

The State of Hawaiʻi exemplifies the challenges of projecting the interactive effects of future climate and land use change on ecosystem carbon balance at a regional scale. Hawaiʻi was the first U.S. state to enact legislation committing to full carbon neutrality, requiring the state to account for and offset all of its greenhouse gas emissions by 2045 (State of Hawaiʻi Acts 15 and 16). This legislation emphasizes the mitigation potential of natural ecosystems as a key component to emissions reduction, necessitating baseline estimates and future projections of land carbon sink strength. However, Hawaiʻi's challenging terrain complicates these assessment efforts. The main Hawaiian Islands are a complex mosaic of natural and human-dominated landscapes overlain by steep climate gradients across relatively short distances, with mean annual temperature ranging from ~4-24° C [@giambelluca_evapotranspiration_2014] and mean annual rainfall ranging from ~180-9500 mm [@giambelluca_online_2013]. Temperatures have risen rapidly in the Hawaiian Islands since the mid 1970s [@giambelluca_secular_2008] and a long-term drying trend has persisted since the early 1920s [@frazier_spatial_2017], resulting in reduced forest biomass and productivity [@barbosa_effects_2017]. These same drying and warming trends have increased the frequency and intensity of wildland fire [@trauernicht_contemporary_2015; @trauernicht_vegetationrainfall_2019] with predictable negative effects on ecosystem carbon balance [@selmants_baseline_2017]. Ecosystem carbon stocks across the main Hawaiian Islands have also been strongly influenced by the legacy of past land-use change [@osher_effect_2003; @asner_high-resolution_2011]. Thousands of hectares of land were deforested beginning in the late 19th century to clear land for sugar plantations and cattle pasture [@cuddihy_alteration_1990]. Since the mid-20th century, much of this agricultural land has been steadily converted to urban areas, commercial forestry plantations, or simply abandoned and colonized by non-native grass species [@suryanata_diversified_2009; @perroy_evolving_2016].
